---
title: "Sectoral_ETFs"
author: "Mandar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE, warning = FALSE,
                      message = FALSE,echo = TRUE, dpi = 360, warning = FALSE,
                      fig.width = 8, fig.height = 6)

```

A stock market sector is a group of stocks that have a lot in common with each other, usually because they are in similar industries. There are 11 different stock market sectors, according to the most commonly used classification system, known as the Global Industry Classification Standard (GICS).
At a glance, the 11 GICS stock market sectors are:

Energy
Materials
Industrials
Utilities
Healthcare
Financials
Consumer Discretionary
Consumer Staples
Information Technology
Communication Services
Real Estate


# For our data on ETFs we focus on
Vanguard sectoral ETFs 
VOX
Communication Services ETF

VCR
Consumer Discretionary ETF

VDC
Consumer Staples ETF

VDE
Energy ETF

VFH
Financials ETF

VHT
Health Care ETF

VIS
Industrials ETF

VGT
Information Technology ETF

VAW
Materials ETF

VNQ
Real Estate ETF

VPU
Utilities ETF



```{r}
# Prequisite libraries
library(tidyquant)
library(tidymodels)
library(tidyverse)
library(tidyfinance)
library(broom)
library(lubridate)
```

# Downloading the data for Vanguard ETFs
```{r}
symbols <- c("VOX","VCR","VDC","VDE","VFH",
             "VHT","VIS","VGT","VAW","VNQ",
             "VPU")

 vang_prices <- tq_get(symbols, get = "stock.prices", from = "2004-01-01")
 
 ## Checking for date consistency:- common date 
 # We find that it is better to use 2005 as the starting
 vang_prices |> 
   group_by(symbol) |> 
   summarise(
     min_date = min(date),
     max_date = max(date)
   )
 #changing the date for consistency

vang_prices <- tq_get(symbols, get = "stock.prices", from = "2005-01-01")
# saving the csv file
write.csv(vang_prices, file = "/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/vang_prices.csv", row.names = FALSE)

#saving as rds file
saveRDS(vang_prices, file = "/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/vang_prices.rds")




```

## Downloading the data for  
Ishares 
ETFs: Sector
IYW: Technology
IYF: Financials
IYR: Real Estate
IYH: Healthcare
IYZ: Communications
IYC: Consumer Discretionary
IYJ: Industrials
IYE: Energy
IDU: Utilites
IYK: Consumer Staples
IYM: Materials 

```{r}
symbols <- c("IYW","IYF","IYR",
"IYH",
"IYZ",
"IYC",
"IYJ",
"IYE",
"IDU",
"IYK",
"IYM")
ishares_prices <- tq_get(symbols, get = "stock.prices", from = "2004-01-01")
 
# date consistency
ishares_prices |> 
   group_by(symbol) |> 
   summarise(
     min_date = min(date),
     max_date = max(date)
   )

saveRDS(ishares_prices, file = "/Users/mandarphatak/Documents/Github/Asset_Pricing/00_Data/ishares_prices.rds")
write_csv(ishares_prices,file = "/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/ishares.csv")
```
# Sectoral Index S&P 500

We also download the data for sectoral index of S&P500 for understanding the performance of the Sectoral ETFs 

1:S&P 500 Materials (Sector) (^SP500-15)
2:S&P 500 Industrials (Sector) (^SP500-20)
3:S&P 500 Financials (Sector) (^SP500-40)
4:S&P 500 Information Technology (^SP500-45)
5:S&P 500 Communication Services (^SP500-50)
6:S&P 500 Consumer Staples (Sector) (^SP500-30)
7:S&P 500 Health Care (Sector) (^SP500-35)
8:S&P 500 Real Estate (Sector) (^SP500-60)
9:S&P 500 Energy (Sector) (^GSPE)
10:S&P 500 Consumer Discretionary (^SP500-25)
11:S&P 500 Utilities (Sector)(^SP500-55)

```{r}
symbols <- c("^SP500-15","^SP500-20","^SP500-40","^SP500-45",
             "^SP500-50","^SP500-30","^SP500-35","^SP500-60",
             "^GSPE","^SP500-25","^SP500-55")
sp500_sector_prices <- tq_get(symbols, get="stock.prices", from ="2000-01-01")
sp500_sector_prices |> 
  group_by(symbol) |> 
  summarise(
    min_date = min(date),
    max_date = max(date)
  )
# common_date is 2002-03-01
sp500_sector_prices <- tq_get(symbols, get="stock.prices", from ="2002-03-01")
# Saving the data

saveRDS(sp500_sector_prices, file = "/Users/mandarphatak/Documents/Github/Asset_Pricing/00_Data/sp500_sectors.rds")
write_csv(sp500_sector_prices,file = "/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/sp500_sectors.csv")
```
# Monthly data 
```{r}
sp500_sector_prices <- readRDS("/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/sp500_sectors.rds")

sp500_sector_rets <- sp500_sector_prices |> 
  group_by(symbol) |> 
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period     = "monthly",
               col_rename = "returns") |> 
  ungroup() |> 
  # for making the data consistent with the dates
  mutate(date = rollback(date, roll_to_first = TRUE))
```

# Saving the data
```{r}
saveRDS(sp500_sector_rets, file = "/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/sp500_sector_rets.rds")
```









