---
title: "Asset_Modeling"
author: "Mandar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE, warning = FALSE,
                      message = FALSE,echo = TRUE, dpi = 360, warning = FALSE,
                      fig.width = 10, fig.height = 8)
```
##First of the Asset Pricing models is the  CAPM model:- 
The capital asset pricing model (CAPM) was developed in the early 1960s by financial economists William Sharpe, Jack Treynor, John Lintner, and Jan Mossin, who built their work on ideas put forth by Harry Markowitz in the 1950s.
CAPM is based on the relationship between an asset’s beta, the risk-free rate (typically the Treasury bill rate), and the equity risk premium, or the expected return on the market minus the risk-free rate.
CAPM evolved as a way to measure this systematic risk. It is widely used throughout finance for pricing risky securities and generating expected returns for assets, given the risk of those assets and cost of capital.
ERi   =  Rf +ßi (ERm−Rf)
where:ERi  = Expected Return of investment
Rf = Risk Free Rate 
ßi =   Beta of the investment
(ERm – Rf) = market risk premium

The risk free rate (Rf) is mostly the treasury rate, in US context
The three-month U.S. Treasury bill is a useful proxy because the market considers there to be virtually no chance of the U.S. government defaulting on its obligations. The large size and deep liquidity of the market contribute to the perception of safety.

Beta compares a security or portfolio's volatility or systematic risk to the market. If a stock is riskier than the market, it will have a beta greater than one. If a stock has a beta of less than one, the formula assumes it will reduce the risk of a portfolio.


#We use the S&P 500 to represent the ERm, we could also use the fama-french single factor  for the same. 
# The risk free rate is taken from the fama-french data. 

CAPM model :-  
```{r}
## Prequisite Library
library(tidyquant)
library(tidyverse)

```

# Step1:- Downloading the data for S&P 500 from yahoo finance
```{r}
sp500 <- tq_get("^GSPC", get = "stock.prices", from = "2004-01-01")
sp500_rets <- sp500 |> 
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "monthly",
               col_rename = "sp500_ret") |> 
  mutate(date = rollback(date,roll_to_first = TRUE))
saveRDS(sp500_rets, file = "/Users/mandarphatak/Documents/Github/Asset_Pricing/00_Data/sp500_rets.rds")
```
# Step 2:- Implementing the CAPM model
```{r}
ishares_rets <- readRDS(file = "/Users/mandarphatak/Documents/Github/Asset_Pricing/00_Data/ishares_rets.rds")
sp500_rets <- readRDS( file = "/Users/mandarphatak/Documents/Github/Asset_Pricing/00_Data/sp500_rets.rds")

data <- ishares_rets |> left_join(sp500_rets, by = "date")

#CAPM model
data %>%
     split(.$symbol) %>% # from base R
    map(~ lm(ishares_ret ~ sp500_ret, data = .)) %>%
    map(summary) 
     
```



