---
title: "Asset_Returns"
author: "Mandar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE, warning = FALSE,
                      message = FALSE,echo = TRUE, dpi = 360, warning = FALSE,
                      fig.width = 10, fig.height = 8)

```


```{r}
# Prequisite libraries
library(tidyquant)
library(tidymodels)
library(tidyverse)
library(tidyfinance)
library(broom)
library(lubridate)

```

# Data manipulation:- Converting the daily data to returns (monthly returns)
```{r}
# reading the data
ishares_prices <- readRDS("/Users/mandarphatak/Documents/GitHub/Asset_Pricing/00_Data/ishares_prices.rds")
# Monthly Returns
ishare_rets <- ishares_prices |> 
  group_by(symbol) |> 
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period     = "monthly",
               col_rename = "returns") |> 
  ungroup() |> 
  # for making the data consistent with the dates
  mutate(date = rollback(date, roll_to_first = TRUE))

```

# Data Visualization:- 
```{r}
# this gives mean returns for each sector ETF but it is not very clear
ishare_rets |> 
  group_by(symbol) |> 
  ggplot(aes(date,returns, colour = symbol)) +
  geom_line() + facet_wrap(~symbol, scales = "free",ncol =4) +
  scale_x_date(
        limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
        date_breaks = "3 years",
        date_labels = "%Y",
        expand = c(0,0)
    ) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  labs(
    title = "Sectoral ETFs Monthly Returns",
    subtitle = "Period:2004-2025",
    caption = "ishares ETFs",
    x = "",
    y = "" 
  ) +
  theme(
       plot.title.position = "plot",
        plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
        plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
        axis.text.y = element_text(face = "bold", color = "black"),
        axis.title = element_text(face = "bold", color = "black"),
        axis.line = element_line(color = "black", linewidth  = 1),
        panel.background = element_rect(fill = "grey", color = NA),
        plot.background = element_rect(fill = "grey", color = NA),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "lightgrey", linewidth  = 0.5),
        axis.line.y.right = element_line(color = "black", linewidth  = 1),
        axis.text.y.right = element_text(face = "bold", color = "black"),
        axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
       legend.position = "none"
        
    ) 

```

## From above we observe that all the sectoral ETFS witnessed the drawdowns during the financial crisis period (2007-2008) and followed by the recession in the US economy. Every ETFs reversed the negative trend in the after the 2009 onwards , this could be attributed to the Fed intervention. In comparison,  Covid (Pandemic)  also led to the underperformance   ETFs, but the extent of negative shock in form of the returns was less compared to that of Financial Crisis.

## So lets dive into each sectoral ETFs and see how each ETF performed during the full time period


### IYW: Technology
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2009-06-30")
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2020-04-30")
covid_final <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYW") |> 
  ggplot(aes(date, returns)) + 
  # Financial Crisis rectangle
  annotate("rect", 
           xmin = recession_start, 
           xmax = recession_end,
           ymin = -Inf, 
           ymax = Inf,
           alpha = 0.2,
           fill = "gray") +
  # COVID recession rectangle
  annotate("rect", 
           xmin = covid_start, 
           xmax = covid_end,
           ymin = -Inf, 
           ymax = Inf,
           alpha = 0.2,
           fill = "gray") +
  # COVID period rectangle
  annotate("rect", 
           xmin = covid_end, 
           xmax = covid_final,
           ymin = -Inf, 
           ymax = Inf,
           alpha = 0.1,
           fill = "darkgrey") +
  
  # Vertical lines
  geom_vline(xintercept = recession_start, 
             linetype = "dashed", 
             color = "darkred", 
             linewidth = 0.5) +
  geom_vline(xintercept = recession_end, 
             linetype = "dashed", 
             color = "darkred", 
             linewidth = 0.5) +
  geom_vline(xintercept = covid_start, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  geom_vline(xintercept = covid_end, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  geom_vline(xintercept = covid_final, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  
  geom_line(col = "blue") +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
  scale_y_continuous(
    limits = c(-.170, .150),
    expand = c(0,0), 
    breaks = c(seq(-.170, .150, .05), .150), 
    labels = function(x) paste0(x, "%")
  ) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  
  # Annotations positioned between lines
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = 0.140,
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 2.5) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = 0.140,
           label = "COVID\nRecession",
           color = "black",
           fontface = "bold",
           size = 2.5) +
  
  annotate("text",
           x = covid_end + (covid_final - covid_end)/2,  # Middle of COVID period
           y = 0.140,
           label = "COVID Period",
           color = "black",
           fontface = "bold",
           size = 2.5) +
  
  labs(
    title = "IYW Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
  
  
```
### IDU Utilites 
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IDU") |> 
  ggplot(aes(date, returns)) + 
  # Financial Crisis rectangle
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694") +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
   scale_y_continuous(
    limits = c(-.130, .120),
    expand = c(0,0), 
    breaks = c(seq(-.130, .120, .05), .120), 
     labels = function(x) sprintf("%.1f%%", x * 100)
  ) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +

  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = 0.09,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = 0.09,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +

  labs(
    title = "IDU Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )

```
### IYF Financials
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYF") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line() +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
  scale_y_continuous(
    limits = c(-.3,.3),
    breaks = seq(-.3,.3,.05), 
    expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .25,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .25,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYF Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
###IYR: Real Estate
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYR") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694") +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
  scale_y_continuous(
    limits = c(-.3,.3),
    breaks = seq(-.3,.3,.05), 
    expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .25,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .25,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYR Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
###IYH: Healthcare
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYH") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694") +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
  scale_y_continuous(
    limits = c(-.15,.15),
    breaks = seq(-.15,.15,.05), 
    expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .1,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .1,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYH Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
### IYZ: Communications

```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYZ") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694",linewidth = 0.7) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
  scale_y_continuous(
    limits = c(-.15,.15),
    breaks = seq(-.15,.15,.05), 
    expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .1,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .1,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYZMonthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
###IYC: Consumer Discretionary
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYC") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694",linewidth = 0.7) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "2 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) + coord_cartesian(ylim = c(-.2,.2)) +
  scale_y_continuous(
#     limits = c(-.3,.15),
   breaks = seq(-.2,.2,.05), 
     expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .15,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .15,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYC Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
### IYJ: Industrials
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYJ") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694",linewidth = 0.7) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "2 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) + coord_cartesian(ylim = c(-.25,.2)) +
  scale_y_continuous(
#     limits = c(-.3,.15),
   breaks = seq(-.25,.2,.05), 
     expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .15,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .15,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYJ Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
### IYE: Energy

```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYE") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694",linewidth = 0.7) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "2 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) + coord_cartesian(ylim = c(-.4,.35)) +
  scale_y_continuous(
#     limits = c(-.3,.15),
   breaks = seq(-.4,.35,.05), 
     expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .25,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .25,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYE Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```

### IYK: Consumer Staples
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYK") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694",linewidth = 0.7) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "2 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) + coord_cartesian(ylim = c(-.2,.15)) +
  scale_y_continuous(
#     limits = c(-.3,.15),
   breaks = seq(-.2,.15,.05), 
     expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .1,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .1,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYK Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```
### IYM : Materials
```{r}
# Define recession dates
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2010-01-01")  # Extended to cover early 2010
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2023-05-01")

ishare_rets |> 
  filter(symbol == "IYM") |> 
  ggplot(aes(date, returns)) + 
  geom_rect(aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.2) +  # Light shading for recession
  # COVID rectangle
  geom_rect(aes(xmin = covid_start, xmax = covid_end, ymin = -Inf, ymax = Inf),
            fill = "#89CFF0", alpha = 0.2) +  # Light shading for COVID
  
  geom_line(col = "#034694",linewidth = 0.7) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "2 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) + coord_cartesian(ylim = c(-.3,.25)) +
  scale_y_continuous(
#     limits = c(-.3,.15),
   breaks = seq(-.3,.25,.05), 
     expand = c(0,0), 
    labels = function(x) sprintf("%.1f%%", x * 100)  # Format to 1 decimal place
) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  # Annotations positioned within plot range
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = .2,  # Adjusted within y-limits
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 3) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = .2,  # Adjusted within y-limits
           label = "COVID",
           color = "black",
           fontface = "bold",
           size = 3) +
  labs(
    title = "IYM Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "ishares ETF",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
```



Performance of S&P500 Tech Sector
#In order to have a more clear idea about the performance of technology sector we use the S&P 500 technology sector index performance:- 

```{r}

# Data for S&P500 Technology Sector Index 
tech_sectors_prices <- tq_get("^SP500-45", get = "stock.prices", from = "2004-01-01")

# Monthly returns
tech_rets <- tech_sectors_prices |> 
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period     = "monthly",
               col_rename = "returns") |> 
  # for making the data consistent with the dates
  mutate(date = rollback(date, roll_to_first = TRUE))


```

# S&P 500 Information Technology (^SP500-45) 
```{r}
recession_start <- as.Date("2007-12-01")
recession_end <- as.Date("2009-06-30")
covid_start <- as.Date("2020-02-01")
covid_end <- as.Date("2020-04-30")
covid_final <- as.Date("2023-05-01")
tech_rets |> 
  ggplot(aes(date, returns)) +
  geom_line(col = "#32a84f", linwidth = 3) +
  # Vertical lines
  geom_vline(xintercept = recession_start, 
             linetype = "dashed", 
             color = "darkred", 
             linewidth = 0.5) +
  geom_vline(xintercept = recession_end, 
             linetype = "dashed", 
             color = "darkred", 
             linewidth = 0.5) +
  geom_vline(xintercept = covid_start, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  geom_vline(xintercept = covid_end, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  geom_vline(xintercept = covid_final, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  scale_x_date(
    limits = c(as.Date("2004-01-01"), as.Date("2025-02-01")),
    date_breaks = "3 years",
    date_labels = "%Y",
    expand = c(0,0)
  ) +
  scale_y_continuous(
    limits = c(-.170, .150),
    expand = c(0,0), 
    breaks = c(seq(-.170, .150, .05), .150), 
    labels = function(x) paste0(x, "%")
  ) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  # Annotations positioned between lines
  annotate("text",
           x = recession_start + (recession_end - recession_start)/2,  # Middle of financial crisis
           y = 0.140,
           label = "Great Recession\n(Financial Crisis)",
           color = "darkred",
           fontface = "bold",
           size = 2.5) +
  
  annotate("text",
           x = covid_start + (covid_end - covid_start)/2,  # Middle of COVID recession
           y = 0.140,
           label = "COVID\nRecession",
           color = "black",
           fontface = "bold",
           size = 2.5) +
  
  annotate("text",
           x = covid_end + (covid_final - covid_end)/2,  # Middle of COVID period
           y = 0.140,
           label = "COVID Period",
           color = "black",
           fontface = "bold",
           size = 2.5) +
  
  labs(
    title = "S&P 500 Information Technology Monthly Returns",
    subtitle = "Period: 2004-2025",
    caption = "S&P",
    x = "",
    y = ""
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0, face = "bold", colour = "black"),
    plot.subtitle = element_text(hjust = 0, face = "bold", colour = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.line = element_line(color = "black", linewidth = 1),
    panel.background = element_rect(fill = "grey", color = NA),
    plot.background = element_rect(fill = "grey", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.5),
    axis.line.y.right = element_line(color = "black", linewidth = 1),
    axis.text.y.right = element_text(face = "bold", color = "black"),
    axis.title.y.right = element_text(face = "bold", color = "black", angle = 90),
    legend.position = "none"
  )
  
  
```


## From above timeseries plot of S&P500 technology sector index we observe similar drawdowns and upvard trends over the perido of time. The range of drawdown is a bit less than the IYW due to being an index.During Covid the performance based on returns is nearly the same.


### Lets see how much investment 

```{r}
library(ggplot2)
library(dplyr)

# Calculate cumulative returns
plot_investment_growth <- function(data) {
  growth_data <- data |>
    mutate(
      cum_return = cumprod(1 + returns),  # Calculate cumulative returns
      investment_value = cum_return       # Assuming $1 initial investment
    )
  
  # Create the plot
  ggplot(growth_data, aes(x = date, y = investment_value)) +
    geom_line(color = "#2c3e50", size = 1) +
    
    labs(
      title = "Growth of $1 Investment in S&P 500 Technology Sector",
      x = "Date",
      y = "Investment Value ($)",
      caption = "Data source: Yahoo Finance"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      axis.title = element_text(size = 12),
      panel.grid.minor = element_blank()
    ) +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y")
}

# Generate the plot
 plot_investment_growth(tech_rets)
```
```{r}
plot_investment_growth <- function(data) {
  growth_data <- data |>
    mutate(
      cum_return = cumprod(1 + returns),  # Calculate cumulative returns
      investment_value = cum_return       # Assuming $1 initial investment
    )
  
  # Create the plot
  ggplot(growth_data, aes(x = date, y = investment_value)) +
    geom_line(color = "#2c3e50", size = 1) +
    labs(
      title = "Growth of $1 Investment in IYW ",
      x = "Date",
      y = "Investment Value ($)",
      caption = "Data source: Yahoo Finance"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      axis.title = element_text(size = 12),
      panel.grid.minor = element_blank()
    ) +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y")
}

# Generate the plot
plot_investment_growth(ishare_rets |> filter(symbol == "IYW"))
```

# 

